Effects of COVID-19 related social media use on well-being

### 2023-07-21 manuscript.tex
- reduced in length

Why are the effects of social media use on well-being small on average?
Two prominent media effect theories argue implicitly against strong average negative effects.
First, according to mood management theory (Zillmann, 1988), using media can affect people's moods.
Use can be stimulating or overwhelming, relaxing or boring.
After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).
Those media that eventually become part of one's media repertoire hence, on average, tend to be beneficial for users to regulate their mood (Marciano et al., 2022).
In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities (Hall \& Liu, 2022).
As with many other things, there can be too much of a good thing.
It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).
However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being (Duradoni et al., 2020).
In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media use represent a new regular behavior that should not be pathologized (Galer, 2018; van Rooij et al., 2018).

@@

If we look at COVID-19 related use more specifically, how could the various types and channels of COVID-19 related social media use affect well-being?
Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
Access to information and resources:
Despite incorrect information, social media provide a vast platform for disseminating accurate and timely information about COVID-19 (John Hopkins University, 2023).
It allows individuals to stay informed about the latest updates, guidelines, and recommendations from reputable health organizations and experts.
Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
Educational opportunities:
Social media can be a valuable educational tool, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
This can include explanatory videos, infographics, and articles that help individuals understand the virus, its transmission, prevention strategies, and vaccination information.
Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
Community support and solidarity:
Social media platforms enable individuals to connect with others who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
Sharing experiences, offering support, and expressing solidarity can foster a sense of community and reduce feelings of isolation.
Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
Promoting mental health and well-being:
Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for maintaining mental well-being during the pandemic (Twitter, 2020).
They provide guidance on coping mechanisms, stress reduction techniques, and self-care practices.
Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
Encouraging positive behaviors:
Social media campaigns and initiatives can promote positive behaviors related to COVID-19, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.

On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
Misinformation and rumors:
Social media platforms can easily spread false or misleading information about COVID-19 (Li et al., 2020).
Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
Overwhelm and anxiety:
Constant exposure to COVID-19-related content on social media can lead to information overload and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
Confirmation bias and echo chambers:
Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to echo chambers (Cinelli et al., 2021).
This might lead to confirmation bias, where individuals are more likely to be exposed to information that confirms their preexisting beliefs.
In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
Cyberbullying and harassment:
Social media platforms are known for fostering negativity, with users sometimes engaging in cyberbullying and harassment (Giumetti \& Kowalski, 2022).
Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
Social comparison and fear of missing out (FOMO):
Social media often showcase the highlights and accomplishments of others, encouraging social comparison (Przybylski et al., 2013).
During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).
This might impact mental health, as most humans have an innate need for social support and physical presence, which online interactions might not fully replicate.

There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
Echoing the theoretical rationales outlines above, studies have yielded mixed results.
Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Klein, 2021; Stainback et al., 2020; Yue et al., 2022).
Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms (Bendau et al., 2021).
Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
On the other hand, some studies reported positive outcomes.
Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being (Guazzini et al., 2022).
Additionally, engaging more on social media was associated with reduced feelings of loneliness (Latikka et al., 2022).
A couple of studies reported mostly neutral effects of social media use on well-being indicators (Bradley \& Howard, 2021; Eden et al., 2020; Sewall et al., 2021).
Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being (see also Bradley \& Howard, 2021; Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021).

@@

The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study.
The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 34 waves.
Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
Panel members were incentivized with 180 credit points for each wave of the study.
The study was conducted between March 2020 and February 2023.
Between March 2020 and July 2020, the intervals between waves were weekly, until May 2022 (wave 32) monthly, and afterward after 5 months.
Each wave consists of at least 1,500 respondents.
Panel mortality was compensated through a continuous re-acquisition of new participants.
The sample size was \emph{N} = 3,641, with overall 123,794 observations.
For an overview of the study set-up, see Figure \ref{fig:study-desc}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-atheyDigitalPublicHealth2023}{}}%
Athey, S., Grabarz, K., Luca, M., \& Wernerfelt, N. (2023). Digital public health interventions at scale: {The} impact of social media advertising on beliefs and outcomes related to {COVID} vaccines. \emph{Proceedings of the National Academy of Sciences}, \emph{120}(5), e2208110120.

\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074.

\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
Bendau, A., Petzold, M. B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., Große, J., Ströhle, A., \& Plag, J. (2021). Associations between {COVID-19} related media consumption and symptoms of anxiety, depression and {COVID-19} related fear in the general population in {Germany}. \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.

\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
Beyens, I., Pouwels, J. L., van Driel, I. I., Keijsers, L., \& Valkenburg, P. M. (2021). Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns. \emph{Communication Research}.

\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. {OSF Preprints}.

\leavevmode\vadjust pre{\hypertarget{ref-cinelliEchoChamberEffect2021}{}}%
Cinelli, M., De Francisci Morales, G., Galeazzi, A., Quattrociocchi, W., \& Starnini, M. (2021). The echo chamber effect on social media. \emph{Proceedings of the National Academy of Sciences}, \emph{118}(9), e2023301118.

\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15.

\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}.

\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142.

\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
Dienlin, T., Masur, P. K., \& Trepte, S. (2017). Displacement or reinforcement? {The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction. \emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.

\leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
Dormann, C., \& Griffin, M. A. (2015). \href{https://www.ncbi.nlm.nih.gov/pubmed/26322999}{Optimal time lags in panel studies}. \emph{Psychological Methods}, \emph{20}(4), 489--505.

\leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T. (2021). \emph{How do good and bad news impact mood during the {Covid-19} pandemic? {The} role of similarity}. {OSF Preprints}.

\leavevmode\vadjust pre{\hypertarget{ref-duradoniWellbeingSocialMedia2020}{}}%
Duradoni, M., Innocenti, F., \& Guazzini, A. (2020). Well-being and social media: {A} systematic review of {Bergen Addiction Scales}. \emph{Future Internet}, \emph{12}(2), 24.

\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
Eden, A. L., Johnson, B. K., Reinecke, L., \& Grady, S. M. (2020). Media for coping during {COVID-19} social distancing: {Stress}, anxiety, and psychological well-being. \emph{Frontiers in Psychology}, \emph{11}, 577639.

\leavevmode\vadjust pre{\hypertarget{ref-egerStatisticalMetaanalysisWellbeing2015}{}}%
Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74.

\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}.

\leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5), 62.

\leavevmode\vadjust pre{\hypertarget{ref-fanStigmatizationSocialMedia2020}{}}%
Fan, L., Yu, H., \& Yin, Z. (2020). Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} -19 on {Twitter}. \emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).

\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
Ferguson, C. J., Kaye, L. K., Branley-Bell, D., Markey, P., Ivory, J. D., Klisanin, D., Elson, M., Smyth, M., Hogg, J. L., McDonnell, D., Nichols, D., Siddiqui, S., Gregerson, M., \& Wilson, J. (2021). Like this meta-analysis: {Screen} media and mental health. \emph{Professional Psychology: Research and Practice}.

\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: {Sense} and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168.

\leavevmode\vadjust pre{\hypertarget{ref-galerHowMuchToo2018}{}}%
Galer, S. S. (2018). \emph{How much is {``too much time''} on social media?} https://www.bbc.com/future/article/20180118-how-much-is-too-much-time-on-social-media.

\leavevmode\vadjust pre{\hypertarget{ref-giumettiCyberbullyingSocialMedia2022}{}}%
Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314.

\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M. (2022). Through the second wave: {Analysis} of the psychological and perceptive changes in the {Italian} population during the {COVID-19} pandemic. \emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.

\leavevmode\vadjust pre{\hypertarget{ref-hallSocialMediaUse2022}{}}%
Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339.

\leavevmode\vadjust pre{\hypertarget{ref-halpernSocialMediaCatalyst2013}{}}%
Halpern, D., \& Gibbs, J. (2013). Social media as a catalyst for online deliberation? {Exploring} the affordances of {Facebook} and {YouTube} for political expression. \emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.

@@

\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
Huang, C. (2017). \href{https://www.ncbi.nlm.nih.gov/pubmed/28622031}{Time spent on social network sites and psychological well-being: {A} meta-analysis}. \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354.

\leavevmode\vadjust pre{\hypertarget{ref-huntSocialMediaBased2022}{}}%
Hunt, I. de V., Dunn, T., Mahoney, M., Chen, M., Nava, V., \& Linos, E. (2022). \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9382165}{A social media-based public health campaign encouraging {COVID-19} vaccination across the {United States}}. \emph{American Journal of Public Health}, \emph{112}(9), 1253--1256.

\leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
John Hopkins University. (2023). {COVID-19 Map}. In \emph{Johns Hopkins Coronavirus Resource Center}. https://coronavirus.jhu.edu/map.html.

\leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509.

\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N. S., Partheymüller, J., Plescia, C., Schiestl, D. W., \& Schlogl, L. (2020). \emph{Austrian {Corona Panel Project} ({SUF} edition)}. {AUSSDA}.

\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N. S., Partheymüller, J., Plescia, C., Schiestl, D. W., \& Schlogl, L. (2021). The {Austrian Corona Panel Project}: Monitoring individual and societal dynamics amidst the {COVID-19} crisis. \emph{European Political Science}, \emph{20}(2), 318--344.

\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling.

\leavevmode\vadjust pre{\hypertarget{ref-lakensEquivalenceTestingPsychological2018}{}}%
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence testing for psychological research: {A} tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269.

>>>>

Why are the effects of social media use on well-being small on average?
Two prominent media effect theories argue implicitly against strong average negative effects.
First, according to mood management theory (Zillmann, 1988), using media can affect people's moods.
Use can be stimulating or overwhelming, relaxing or boring.
After some time, users implicitly learn which media help them balance their mood and affect according to their own situational needs (Zillmann, 1988).
Those media that eventually become part of one's media repertoire hence, on average, tend to be beneficial for users to regulate their mood (Marciano et al., 2022).
In conclusion, if a certain medium is used frequently, mood-management theory argues that it is likely not detrimental for well-being.
On the other hand, although mood management theory suggests that effects should not be overly negative, it could be that although short-term effects are positive, long-term effects are negative.
Precisely because social media have so many positive consequences in the short run, this might cause problem in the long run---for example, because social media use displaces other potentially more meaningful and/or relaxing activities (Hall \& Liu, 2022).
As with many other things, there can be too much of a good thing.
It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves (Yang et al., 2021).

@@

If we look at COVID-19 related use more specifically, how could the various types and channels of COVID-19 related social media use affect well-being?
Several uses and gratifications exist, which help explain why people used social media frequently during the pandemic.
Despite incorrect information, social media provide a vast platform for disseminating \emph{accurate and timely information} about COVID-19 (John Hopkins University, 2023).
Access to reliable information can help people make informed decisions, alleviate uncertainties, and feel empowered during the pandemic.
Social media can be a valuable \emph{educational tool}, with various organizations and experts sharing informative content about COVID-19 (World Health Organization, 2023).
Access to educational resources on social media can empower individuals to make informed decisions regarding their health.
Social media platforms enable individuals to \emph{connect with others} who are experiencing similar challenges during the pandemic (Guazzini et al., 2022).
Engaging in online communities and support groups can provide emotional support and create a network of like-minded individuals.
Many mental health organizations and professionals utilize social media to share tips, strategies, and resources for \emph{maintaining mental well-being} during the pandemic (Twitter, 2020).
Engaging with such content might help individuals prioritize their mental health and develop resilience during challenging times.
Social media campaigns and initiatives can promote \emph{positive COVID-19 behaviors}, such as mask-wearing, physical distancing, hand hygiene, and vaccination (Athey et al., 2023; Hunt et al., 2022).
Public health organizations and influencers leverage the power of social media to spread awareness and encourage responsible actions, contributing to public health efforts and fostering a sense of collective responsibility.

On the other hand, the effects might be negative, perhaps best explained by the following five mechanisms.
Social media platforms can easily spread \emph{false or misleading information} about COVID-19 (Li et al., 2020).
Due to the ease of sharing and the lack of fact-checking, inaccurate information can go viral and might cause confusion, anxiety, and panic among users.
Constant exposure to COVID-19-related content on social media can lead to \emph{information overload} and contribute to heightened anxiety levels (J. Fan \& Smith, 2021).
The rapid spread of news, updates, and opinions can be overwhelming and might exacerbate existing stress or fears about the pandemic (Sharma et al., 2022).
Social media algorithms are designed to show users content that aligns with their interests and beliefs, potentially contributing to \emph{echo chambers} (Cinelli et al., 2021).
In the context of COVID-19, this might reinforce false or misleading information and prevent individuals from considering alternative perspectives.
Social media platforms are known for fostering negativity, with users sometimes engaging in \emph{cyberbullying and harassment} (Giumetti \& Kowalski, 2022).
Discussions around COVID-19 can become heated and polarized, leading to personal attacks and online conflicts.
Such experiences threaten mental well-being and might contribute to feelings of distress and isolation.
Social media often showcase the highlights and accomplishments of others, encouraging \emph{social comparison} (Przybylski et al., 2013).
During a pandemic, seeing posts about others' successes or seemingly perfect lives might intensify feelings of inadequacy or FOMO, especially when individuals are unable to participate in similar activities due to restrictions or personal circumstances (Sharma et al., 2022).

There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
Echoing the theoretical rationales outlines above, studies have yielded mixed results.
Some studies found negative effects, indicating that excessive social media use for COVID-19 news led to compulsive behavior and increased stress levels, particularly due to upward social comparison (Klein, 2021; Stainback et al., 2020; Yue et al., 2022).
Individuals who relied on social media as their primary information source reported higher levels of anxiety and depression symptoms (Bendau et al., 2021).
Similarly, increased COVID-19-related media consumption was associated with higher psychological distress.
On the other hand, some studies reported positive outcomes.
Certain individuals experienced increased virtual community and social connectedness during the pandemic through social media, which contributed to their well-being (Guazzini et al., 2022).
Additionally, engaging more on social media was associated with reduced feelings of loneliness (Latikka et al., 2022).
A couple of studies reported mostly neutral effects of social media use on well-being indicators (Bradley \& Howard, 2021; Eden et al., 2020; Sewall et al., 2021).
Overall, the literature demonstrates a mixed picture, highlighting both positive and negative effects of social media use focused on or during COVID-19 on well-being.

@@

The data come from the Austrian Corona Panel Project (Kittel et al., 2020), which is a large-scale standalone panel study.
The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 34 waves.
Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
Panel members were incentivized with 180 credit points for each wave of the study.
The study was conducted between March 2020 and February 2023.
Between March 2020 and July 2020, the intervals between waves were weekly, until May 2022 (wave 32) monthly, and afterward after 5 months.
Each wave consists of at least 1,500 respondents.
Panel mortality was compensated through a continuous re-acquisition of new participants.
The sample size was \emph{N} = 3,641, with overall 123,794 observations.
For an overview of the study set-up, see Figure \ref{fig:study-desc}.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-atheyDigitalPublicHealth2023}{}}%
Athey, S., Grabarz, K., Luca, M., \& Wernerfelt, N. (2023). Digital public health interventions at scale: {The} impact of social media advertising on beliefs and outcomes related to {COVID} vaccines. \emph{Proceedings of the National Academy of Sciences}, \emph{120}(5), e2208110120. \url{https://doi.org/10.1073/pnas.2208110120}

\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074. \url{https://doi.org/10.1007/s11135-018-0802-x}

\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
Bendau, A., Petzold, M. B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., Große, J., Ströhle, A., \& Plag, J. (2021). Associations between {COVID-19} related media consumption and symptoms of anxiety, depression and {COVID-19} related fear in the general population in {Germany}. \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291. \url{https://doi.org/10.1007/s00406-020-01171-6}

\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
Beyens, I., Pouwels, J. L., van Driel, I. I., Keijsers, L., \& Valkenburg, P. M. (2021). Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns. \emph{Communication Research}. \url{https://doi.org/10.1177/00936502211038196}

\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}

\leavevmode\vadjust pre{\hypertarget{ref-cinelliEchoChamberEffect2021}{}}%
Cinelli, M., De Francisci Morales, G., Galeazzi, A., Quattrociocchi, W., \& Starnini, M. (2021). The echo chamber effect on social media. \emph{Proceedings of the National Academy of Sciences}, \emph{118}(9), e2023301118. \url{https://doi.org/10.1073/pnas.2023301118}

\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15. \url{https://doi.org/10.1525/collabra.115}

\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}

\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142. \url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}

\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
Dienlin, T., Masur, P. K., \& Trepte, S. (2017). Displacement or reinforcement? {The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction. \emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87. \url{https://doi.org/10.1111/jcc4.12183}

\leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}

\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
Eden, A. L., Johnson, B. K., Reinecke, L., \& Grady, S. M. (2020). Media for coping during {COVID-19} social distancing: {Stress}, anxiety, and psychological well-being. \emph{Frontiers in Psychology}, \emph{11}, 577639. \url{https://doi.org/10.3389/fpsyg.2020.577639}

\leavevmode\vadjust pre{\hypertarget{ref-egerStatisticalMetaanalysisWellbeing2015}{}}%
Eger, R. J., \& Maridal, J. H. (2015). A statistical meta-analysis of the wellbeing literature. \emph{International Journal of Wellbeing}, \emph{5}(2), 45--74. \url{https://doi.org/10.5502/ijw.v5i2.4}

\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-fanInformationOverloadWellbeing2021}{}}%
Fan, J., \& Smith, A. P. (2021). Information overload, wellbeing and {COVID-19}: {A} survey in {China}. \emph{Behavioral Sciences}, \emph{11}(5, 5), 62. \url{https://doi.org/10.3390/bs11050062}

\leavevmode\vadjust pre{\hypertarget{ref-fanStigmatizationSocialMedia2020}{}}%
Fan, L., Yu, H., \& Yin, Z. (2020). Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}. \emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1). \url{https://doi.org/10.1002/pra2.313}

\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
Ferguson, C. J., Kaye, L. K., Branley-Bell, D., Markey, P., Ivory, J. D., Klisanin, D., Elson, M., Smyth, M., Hogg, J. L., McDonnell, D., Nichols, D., Siddiqui, S., Gregerson, M., \& Wilson, J. (2021). Like this meta-analysis: {Screen} media and mental health. \emph{Professional Psychology: Research and Practice}. \url{https://doi.org/10.1037/pro0000426}

\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: {Sense} and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168. \url{https://doi.org/10.1177/2515245919847202}

\leavevmode\vadjust pre{\hypertarget{ref-giumettiCyberbullyingSocialMedia2022}{}}%
Giumetti, G. W., \& Kowalski, R. M. (2022). Cyberbullying via social media and well-being. \emph{Current Opinion in Psychology}, \emph{45}, 101314. \url{https://doi.org/10.1016/j.copsyc.2022.101314}

\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M. (2022). Through the second wave: {Analysis} of the psychological and perceptive changes in the {Italian} population during the {COVID-19} pandemic. \emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635. \url{https://doi.org/10.3390/ijerph19031635}

\leavevmode\vadjust pre{\hypertarget{ref-hallSocialMediaUse2022}{}}%
Hall, J. A., \& Liu, D. (2022). Social media use, social displacement, and well-being. \emph{Current Opinion in Psychology}, \emph{46}, 101339. \url{https://doi.org/10.1016/j.copsyc.2022.101339}

\leavevmode\vadjust pre{\hypertarget{ref-halpernSocialMediaCatalyst2013}{}}%
Halpern, D., \& Gibbs, J. (2013). Social media as a catalyst for online deliberation? {Exploring} the affordances of {Facebook} and {YouTube} for political expression. \emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168. \url{https://doi.org/10.1016/j.chb.2012.10.008}

@@

\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
Huang, C. (2017). Time spent on social network sites and psychological well-being: {A} meta-analysis. \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354. \url{https://doi.org/10.1089/cyber.2016.0758}

\leavevmode\vadjust pre{\hypertarget{ref-huntSocialMediaBased2022}{}}%
Hunt, I. de V., Dunn, T., Mahoney, M., Chen, M., Nava, V., \& Linos, E. (2022). A social media‒based public health campaign encouraging {COVID-19} vaccination across the {United States}. \emph{American Journal of Public Health}, \emph{112}(9), 1253--1256. \url{https://doi.org/10.2105/AJPH.2022.306934}

\leavevmode\vadjust pre{\hypertarget{ref-johnhopkinsuniversityCOVID19Map2023}{}}%
John Hopkins University. (2023). \emph{{COVID-19 Map}}. {Johns Hopkins Coronavirus Resource Center}. \url{https://coronavirus.jhu.edu/map.html}

\leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}

\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N. S., Partheymüller, J., Plescia, C., Schiestl, D. W., \& Schlogl, L. (2020). \emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}. {AUSSDA}. \url{https://doi.org/10.11587/28KQNS}

\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}

\leavevmode\vadjust pre{\hypertarget{ref-lakensEquivalenceTestingPsychological2018}{}}%
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence testing for psychological research: {A} tutorial. \emph{Advances in Methods and Practices in Psychological Science}, \emph{1}(2), 259--269. \url{https://doi.org/10.1177/2515245918770963}

\leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
Latikka, R., Koivula, A., Oksa, R., Savela, N., \& Oksanen, A. (2022). Loneliness and psychological distress before and during the {COVID-19} pandemic: {Relationships} with social media identity bubbles. \emph{Social Science \& Medicine}, \emph{293}, 114674. \url{https://doi.org/10.1016/j.socscimed.2021.114674}
